\Huge \textbf{Hybrid rendering}\\ \huge using Radeon Rays

### 2019-05-27 Paper/Paper_BillietStijn_RadeonRays.tex
Improved hard shadow implementation in awaition for soft shadow changes

\begin{figure}[h]
	\caption{Illustration of different ray-casting solutions}
	\centering
	\includegraphics[width=0.95\textwidth]{raycasting.jpg}
\end{figure}
\subsubsection{Definition}
In short ray-tracing is a rendering technique that can produce lifelike lighting effects. A ray-tracing algorithm essentially achieves this by sending out rays for every pixel of our framebuffer (camera) towards the corresponding scene geometry. Based on the technique we can send out additional bounces (secondary rays). It then basically backtracks that chain of light-rays to accumulate material/light information to then calculate the final pixel color.

@@

\begin{figure}[h]
\includegraphics[width=\textwidth]{rasterization.jpg}
\caption{Rasterization graphics pipeline.}
\end{figure}

@@

\subsubsection{Graphics pipeline}
With rasterization objects on screen are created from meshes. A mesh is a collection of three-dimensional points a.k.a. vertices. These vertices are the main resource needed for rendering, these go through what's called the rendering pipeline to then be converted into screen pixels. The rendering pipeline or more formerly known as the graphics pipeline acts as a conceptual model that describes what steps a graphics system needs to performs to be able to render aforementioned meshes to a 2D screen. \newline\par
\pagebreak

\subsubsection{Vertex shader}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{coordinate_systems.png}
	\caption{Model space to screen space.}
	\label{fig:coordinates}
\end{figure}
The first stage is the vertex shader which essentially transforms the vertices to their correct
positions, the vertices have to be transformed seeing that they are defined in their own local transform space(vertices positions are defined relative to their own center point). The vertices are defined this way as a means to make it easier for modelers, they quite simply have to define the vertices relative to the modeling software's gizmo.

\subsubsection{Shape assembly}
\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{primitives.png}
	\caption{List of possible OpenGL primitives.}
\end{figure}

@@

\subsubsection{Geometry shader}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.7\textwidth]{geometryshader.jpg}
	\caption{Purpose of geometry shader.}
\end{figure}
The third stage is an optional geometry shader, this stage won't be covered in depth, but it is essentially in charge of creating additional geometry (extra vertices) at runtime.\newline\par

@@

\begin{minipage}{0.5\textwidth}
	\centering
	\includegraphics[width=\textwidth]{culling.png}
\end{minipage} \hfill
\begin{minipage}{0.5\textwidth}
	The first task of the rasterizer is primitive culling. The shapes that made it to the rasterizer are checked based on their normals. The normal of a primitive/vertex tells us if the face is facing away from us or not. This simple operation can be demonstrated using the dotproduct, however this does get handled by the specifications (e.g. DirectX/OpenGL) rasterizer. In its essence the calculation will translate to if the dotproduct is negative then the face is facing away, if it's positive then it's facing toward us.
\end{minipage}

@@

\begin{minipage}{0.4\textwidth}
	\includegraphics[width=\textwidth]{frustrum.png}
\end{minipage} \hfill
\begin{minipage}{0.59\textwidth}
	The third task of the rasterizer is the homogeneous divide. The previous stages were in charge of transforming the vertices from model space into clipping space, this is done by multiplying the world, view and projection matrix with the vertices this yields a valid projection but does not account for depth information (f.e. objects that are nearby will appear the same as objects from afar) this is where the z-component comes in, seeing the z stores our depth information. Before we apply the projectionmatrix we store the z-component as an additional w component.
\end{minipage}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=0.85\textwidth]{Zbuffer.png}
\end{figure}
The fifth stage, the fragment stage is a shader stage that will process all pixel fragments processed by the rasterizer. Depending on the technique we collapse all information into a set of colors (color buffer) and a single depth value (depth buffer), additionally we can augment this by writing away a stencil value (stencil buffer).

\subsubsection{Test and blending stage}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{alpha.png}
\end{figure}
Finally we've got the testing and blending stage. Up until this point "pixels" where addressed as pixel fragments, this is in-fact a clear distinction to make seeing that pixel fragments are not guaranteed to show up on screen, they could be discarded in the shaders or they could be transparent. If there are transparent objects in the scene we have to keep both pixels, those from the object behind it and the transparent object, we can then simply blend those pixels together to receive the final pixel colour. This blending operation is in-fact quite costly seeing we need to make sure the objects appear in the right order and we've got the load of additional pixels on our hands. A more optimized technique can be used for objects that are fully transparent however, here we can employ alpha testing where we simply discard those pixel fragments in the fragment shader.

@@

\subsection{Raytracing}
\subsubsection{OpenCL}
Open Computing Language is a framework for writing programs that execute across compute devices. This includes among other hardware accelerators, the central processing unit (CPU) and the graphics processing unit (GPU). These aforementioned programs are called kernels and are also written on a pseudo c-language. OpenCL provides the programmer of a standard interface to perform parallel computing tasks. The programmer has to set up the OpenCL execution pipeline and write the kernels.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.9\textwidth]{openclprogram.png}
\end{figure}
Under this section you'll find a quick rundown of aforementioned execution 
model.
\pagebreak
\begin{itemize}
	\item{\textbf{Platform}}\newline\par
	The platform that we are running this code on, so which vendor (e.g. Nvidia).
	\item{\textbf{Device}}\newline\par
	What device are we going to run this on could be anything from CPU, GPU to APU.
	\item{\textbf{Context}}\newline\par
	This is the runtime interface between the host program (run on the CPU) and the device (CPU/GPU/etc.). The context is in charge of managing all the OpenCL resources so programs, kernels, command queues and buffers.
	\item{\textbf{Program}}\newline\par
	The entire OpenCL program, this could be one or more kernels/device functions bundled together into a single program file.
	\item{\textbf{Kernel}}\newline\par
	The starting point/building blocks of a OpenCL program. Kernels are called from the CPU, they provide the basic units of executable code that run on a OpenCL device. The kernels are preceded with the keyword "\_\_kernel"
	\item{\textbf{Command queue}}\newline\par
	The command queue allows the kernels execution commands to be sent to the device. This effectively queues them for execution, this en-queuing can be in-order or out-of-order.
	\item{\textbf{Memory objects}}\newline\par
	The memory objects are what is provided to the kernels. Kernels do accept basic types like floats, ints, bools etc. and self defined types like Ray, Intersection etc, but if there is a large buffer of those it's kept in a Memory buffer (e.g. images).
\end{itemize}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=0.7\textwidth]{memorymodel.png}
	\caption{Illustration of the memory execution model.}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{sharegroups.png}
	\caption{OpenCL/OpenGL interop context}
\end{figure}

@@

	\begin{figure}[h]
		\centering
		\includegraphics[width=0.9\textwidth]{assimpstructure.png}
	\end{figure}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{fvsdef.png}
	\caption{Difference between forward and deferred pipeline.}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{fvsdefgraph.png}
	\caption{Lighting performance comparison between forward and deferred.}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=0.6\textwidth]{ndc.png}
	\caption{Normalized device coordinates for our triangle}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{vertexattributepointer.png}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=0.7\textwidth]{formula.png}
\end{figure}

\begin{minipage}{0.4\textwidth}
	This sampling was needed to try and find out how large the penumbra (shadow gradient) should be. Here it was a simple case of looking for the nearest non occluded (lit) pixel. This required a grid search to find out where in that sample that nearest pixel was.
\end{minipage} \hfill
\begin{minipage}{0.6\textwidth}
	\centering
	\includegraphics[width=0.8\textwidth]{sampler.png}
\end{minipage}

@@

>>>>

\begin{figure}[h]
	\caption{Illustration of different ray-casting solutions \cite{keepmaxing16}}
	\centering
	\includegraphics[width=0.95\textwidth]{raycasting.jpg}
\end{figure}
\subsubsection{Definition}
In short ray-tracing is a rendering technique that can produce lifelike lighting effects. A ray-tracing algorithm essentially achieves this by sending out rays for every pixel of our framebuffer (camera) towards the corresponding scene geometry. Based on the technique we can send out additional bounces (secondary rays). It then basically backtracks that chain of light-rays to accumulate material/light information to then calculate the final pixel color.

@@

\begin{figure}[h]
\includegraphics[width=\textwidth]{rasterization.jpg}
\caption{Rasterization graphics pipeline. \cite{learnopengl18}}
\end{figure}

@@

\subsubsection{Graphics pipeline}\cite{pipeline18}
With rasterization objects on screen are created from meshes. A mesh is a collection of three-dimensional points a.k.a. vertices. These vertices are the main resource needed for rendering, these go through what's called the rendering pipeline to then be converted into screen pixels. The rendering pipeline or more formerly known as the graphics pipeline acts as a conceptual model that describes what steps a graphics system needs to performs to be able to render aforementioned meshes to a 2D screen. \newline\par
\pagebreak

\subsubsection{Vertex shader}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{coordinate_systems.png}
	\caption{Model space to screen space. \cite{learnopengl18}}
	\label{fig:coordinates}
\end{figure}
The first stage is the vertex shader which essentially transforms the vertices to their correct
positions, the vertices have to be transformed seeing that they are defined in their own local transform space(vertices positions are defined relative to their own center point). The vertices are defined this way as a means to make it easier for modelers, they quite simply have to define the vertices relative to the modeling software's gizmo.

\subsubsection{Shape assembly}
\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{primitives.png}
	\caption{List of possible OpenGL primitives. \cite{ratcliffe05}}
\end{figure}

@@

\subsubsection{Geometry shader}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.7\textwidth]{geometryshader.jpg}
	\caption{Purpose of geometry shader. \cite{rastergrid00}}
\end{figure}
The third stage is an optional geometry shader, this stage won't be covered in depth, but it is essentially in charge of creating additional geometry (extra vertices) at runtime.\newline\par

@@

\begin{minipage}{0.5\textwidth}
	\centering
	\includegraphics[width=\textwidth]{culling.png}
	\cite{mayer18}
\end{minipage} \hfill
\begin{minipage}{0.5\textwidth}
	The first task of the rasterizer is primitive culling. The shapes that made it to the rasterizer are checked based on their normals. The normal of a primitive/vertex tells us if the face is facing away from us or not. This simple operation can be demonstrated using the dotproduct, however this does get handled by the specifications (e.g. DirectX/OpenGL) rasterizer. In its essence the calculation will translate to if the dotproduct is negative then the face is facing away, if it's positive then it's facing toward us.
\end{minipage}

@@

\begin{minipage}{0.4\textwidth}
	\includegraphics[width=\textwidth]{frustrum.png}
	\cite{frustrumyourdictionary18}
\end{minipage} \hfill
\begin{minipage}{0.59\textwidth}
	The third task of the rasterizer is the homogeneous divide. The previous stages were in charge of transforming the vertices from model space into clipping space, this is done by multiplying the world, view and projection matrix with the vertices this yields a valid projection but does not account for depth information (f.e. objects that are nearby will appear the same as objects from afar) this is where the z-component comes in, seeing the z stores our depth information. Before we apply the projectionmatrix we store the z-component as an additional w component.
\end{minipage}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=0.85\textwidth]{Zbuffer.png}\cite{Zbuffering18}
\end{figure}
The fifth stage, the fragment stage is a shader stage that will process all pixel fragments processed by the rasterizer. Depending on the technique we collapse all information into a set of colors (color buffer) and a single depth value (depth buffer), additionally we can augment this by writing away a stencil value (stencil buffer).

\subsubsection{Test and blending stage}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{alpha.png}\cite{yourdictionary18}
\end{figure}
Finally we've got the testing and blending stage. Up until this point "pixels" where addressed as pixel fragments, this is in-fact a clear distinction to make seeing that pixel fragments are not guaranteed to show up on screen, they could be discarded in the shaders or they could be transparent. If there are transparent objects in the scene we have to keep both pixels, those from the object behind it and the transparent object, we can then simply blend those pixels together to receive the final pixel colour. This blending operation is in-fact quite costly seeing we need to make sure the objects appear in the right order and we've got the load of additional pixels on our hands. A more optimized technique can be used for objects that are fully transparent however, here we can employ alpha testing where we simply discard those pixel fragments in the fragment shader.

@@

\subsection{Raytracing}
\subsubsection{OpenCL}\cite{lapere16}
Open Computing Language is a framework for writing programs that execute across compute devices. This includes among other hardware accelerators, the central processing unit (CPU) and the graphics processing unit (GPU). These aforementioned programs are called kernels and are also written on a pseudo c-language. OpenCL provides the programmer of a standard interface to perform parallel computing tasks. The programmer has to set up the OpenCL execution pipeline and write the kernels.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.9\textwidth]{openclprogram.png}
\end{figure}
Under this section you'll find a quick rundown of aforementioned execution 
model.\cite{lapere16}
\pagebreak
\begin{itemize}
	\item{\textbf{Platform}}\newline\par
	The platform that we are running this code on, so which vendor (e.g. Nvidia).
	\item{\textbf{Device}}\newline\par
	What device are we going to run this on could be anything from CPU, GPU to APU.
	\item{\textbf{Context}}\newline\par
	This is the runtime interface between the host program (run on the CPU) and the device (CPU/GPU/etc.). The context is in charge of managing all the OpenCL resources so programs, kernels, command queues and buffers.
	\item{\textbf{Program}}\newline\par
	The entire OpenCL program, this could be one or more kernels/device functions bundled together into a single program file.
	\item{\textbf{Kernel}}\newline\par
	The starting point/building blocks of a OpenCL program. Kernels are called from the CPU, they provide the basic units of executable code that run on a OpenCL device. The kernels are preceded with the keyword "\_\_kernel"
	\item{\textbf{Command queue}}\newline\par
	The command queue allows the kernels execution commands to be sent to the device. This effectively queues them for execution, this en-queuing can be in-order or out-of-order.
	\item{\textbf{Memory objects}}\newline\par
	The memory objects are what is provided to the kernels. Kernels do accept basic types like floats, ints, bools etc. and self defined types like Ray, Intersection etc, but if there is a large buffer of those it's kept in a Memory buffer (e.g. images).
\end{itemize}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=0.7\textwidth]{memorymodel.png}
	\caption{Illustration of the memory execution model. \cite{lapere16}}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{sharegroups.png}
	\caption{OpenCL/OpenGL interop context \cite{apple18}}
\end{figure}

@@

	\begin{figure}[h]
		\centering
		\includegraphics[width=0.9\textwidth]{assimpstructure.png}
		\caption{Illustation of AssImp model structure. \cite{learnopengl18}}
	\end{figure}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{fvsdef.png}
	\caption{Difference between forward and deferred pipeline. \cite{owens18}}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{fvsdefgraph.png}
	\caption{Lighting performance comparison between forward and deferred. \cite{vanoosten18}}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=0.6\textwidth]{ndc.png}
	\caption{Normalized device coordinates for our triangle. \cite{learnopengl18}}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{vertexattributepointer.png}\cite{learnopengl18}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=0.7\textwidth]{formula.png}\cite{kissner16}
\end{figure}

\begin{minipage}{0.4\textwidth}
	This sampling was needed to try and find out how large the penumbra (shadow gradient) should be. Here it was a simple case of looking for the nearest non occluded (lit) pixel. This required a grid search to find out where in that sample that nearest pixel was.
\end{minipage} \hfill
\begin{minipage}{0.6\textwidth}
	\centering
	\includegraphics[width=0.8\textwidth]{sampler.png}\cite{kissner16}
\end{minipage}

@@

	\bibitem{keepmaxing16}
	Keep Maxing,
	2016,
	Mastering Vray Lighting \& Materials - 052 Understanding Vray Render Optimization,
	\url{https://www.youtube.com/watch?v=A3z9v2MVRvw}

	\bibitem{learnopengl18}
	Learn OpenGL,
	2018,
	Learn OpenGL,
	\url{https://learnopengl.com/}

	\bibitem{yourdictionary18}
	yourdictionary
	2018,
	alpha-blending,
	\url{http://www.yourdictionary.com/alpha-blending}

	\bibitem{Zbuffering18}
	WikiPedia
	2018,
	Z-buffering,
	\url{https://en.wikipedia.org/wiki/Z-buffering}

	\bibitem{ratcliffe05}
	Stephen Ratcliffe
	2005,
	Review of OpenGL,
	\url{http://vision-simulator.tripod.com/report/section_7.htm}

	\bibitem{rastergrid00}
	RasterGrid
	2000,
	history-of-hardware-tessellation,
	\url{http://rastergrid.com/blog/2010/09/history-of-hardware-tessellation/}

	\bibitem{mayer18}
	A. Julian Mayer,
	2018,
	Dynamic Occlusion Culling,
	\url{https://pdfs.semanticscholar.org/5cab/0fac256b81b71c8d85a51881f7daf52ab7f3.pdf}

	\bibitem{frustrumyourdictionary18}
	yourdictionary
	2018,
	view frustrum,
	\url{http://www.yourdictionary.com/viewing-frustum}

	\bibitem{lapere16}
	Sam Lapera,
	2016,
	OpenCL path tracing tutorial 1: Firing up OpenCL
	\url{http://raytracey.blogspot.com/2016/11/opencl-path-tracing-tutorial-1-firing.html}

	\bibitem{apple18}
	Apple,
	2018,
	Sharing Data Between OpenCL and OpenGL
	\url{https://developer.apple.com/library/archive/documentation/Performance/Conceptual/OpenCL_MacProgGuide/shareGroups/shareGroups.html}

	\bibitem{owens18}
	Ben Owens,
	2018,
	Forward rendering vs deferred
	\url{https://gamedevelopment.tutsplus.com/articles/forward-rendering-vs-deferred-rendering--gamedev-12342}

	\bibitem{vanoosten18}
	Jeremiah van Oosten,
	2018,
	Forward vs Deferred vs Forward+ Rendering with DirectX 11
	\url{https://www.3dgep.com/forward-plus/}
